 

### 2023-03-11 paper2.Rmd
Revisions throughout and final version for External Reader

        val <- round(exp(output_df$estimate), 3)
        p_val <- round(output_df$p_value, 3)
        low <- round(exp(output_df$lower_99), 3)
        high <- round(exp(output_df$upper_99), 3)

@@

          return(paste0(
            "(IRR: ",
            val,
            ", p ",
            ifelse(p_val < 0.001,
                   "< 0.0001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

@@

          return(paste0(
            "(p ",
            ifelse(p_val < 0.001,
                   "< 0.0001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

@@

          return(paste0(ifelse(
            val < 1,
            paste0(round(1 - val, 3) * 100, "\\% fewer"),
            paste0(round(val - 1, 3) * 100, "\\%")
          )))

@@

      if (value == "ame"){
        val <- round(output_df$ame, 3)
        p_val <- round(output_df$ame_p_value, 3)
        low <- round(output_df$ame_lower_99, 3)
        high <- round(output_df$ame_upper_99, 3)

        if (value_include == TRUE){
          return(paste0(
            "(AME: ",
            val,
            ", p ",
            ifelse(p_val < 0.001,
                   "< 0.001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

@@

          return(paste0(
            "(p ",
            ifelse(p_val < 0.001,
                   "< 0.001, ",
                   paste0("= ",
                          p_val,
                          ", ")),
            "99\\% CI: ",
            low,
            " – ",
            high
            ,
            ")"))

@@

          return(round(val, 2))

@@

\emph{Daniel Arias, Karestan Koenen, Christopher Golden}

\vspace{24pt}

@@

{\bf Methods:} To better understand climate conditions and population mental health in these settings, we leverage meteorologic, geospatial, and monthly cases of mental and neurological disorders reported by `r num_facilities_pretty` facilities in Madagascar from 2010 to 2020 to explore the mental health impacts of three CCEs using negative binomial regression: temperature, soil moisture, and duration of cyclones and tropical storms.

{\bf Findings and interpretation:} We found a statistically significant reduction in monthly cases reported associated with temperature—approximately `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = "only")` cases reported per each degree Celsius increase `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = TRUE)`. A multiverse analysis of `r num_models_multiverse` models found this estimate to be robust to alternative specifications; ecoregion analysis indicated that this association was primarily localized to facilities in the central highlands, where cooler temperatures are more prevalent. We further found that higher soil moisture may be associated with higher case counts in the future, with an `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")`  elevation in incident cases reported `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)`  after a three month lag. Ecoloregion analysis suggest higher cases are reported in the eastern lowlands and the northern dry deciduous forests when soil moisture rises, which is consistent with the geographic concentration of flood risk in Madagascar. Our findings highlight the complex interactions between CCEs and population mental health and suggest potential heterogeneity in how changing climate conditions will impact mental health needs and system responses.

@@

Understanding how both incremental and extreme changes in climate impact population mental health is critical to gauging the full extent to which CCEs impact human health and wellbeing under a multidisciplinary approach to population health.[@berryCaseSystemsThinking2018] This understanding is particularly important in contexts such as Madagascar, where climate conditions are already demonstrating pronounced pressures on economic, social, and physical wellbeing.

In investigating the impact of climate exposures on mental health, health system data can serve as a useful tool in settings where epidemiological survey data is sparse. Despite the poor availability of mental health services in many settings, routine health system data can still provide valuable insights into trends and patterns of service use and the need and distribution of care for mental health issues. Although these data may only capture a share of the true burden of disease, reported cases and visits related can be a valuable starting point for preliminary research into the impact of climate change on population mental health in resource-constrained settings.

@@

As a measure of population mental health, we considered facility-level monthly reports cases of mental disorders; in terms of exposure, we explored the predictive association of three CCEs with incident mental health visits: mean monthly ambient air temperature, mean monthly soil moisture, and monthly cyclone activity. As exposure to these stressors could involve a delay in onset of mental health symptoms and subsequent care seeking, we further explored whether exposures in prior months were associated with monthly reported cases in sequent ones.  We further considered interactive and additive models to account for complexity in the interaction of climate variables in Madagascar, as well as ecological variation to capture differences in regional climate systems.

@@

To analyze the impact of climate on population mental health in Madagascar, we first obtained health management information system (HMIS) data from the Madagascar Ministère de la Santé Publique. We then obtained gridded temperature and soil moisture data from the European ReAnalysis (ERA5)[@copernicusclimatechangeservicec3sFifthGenerationECMWF2017] and the European Space Agency’s Climate Change Initiative (ESA CCI),[@europeanspaceagencyEuropeanSpaceAgency2022] respectively. Data on tropical storms and cyclones were obtained from open-source compilations of storms in the South-West Indian Ocean[@wikipediaSouthWestIndianOcean2023], which were in turn cross-referenced against storm trajectory data from the National Oceanic and Atmospheric Administration (NOAA)[@InternationalBestTrack2021] for accuracy.

@@

Exploratory data analysis indicated that the classification and reporting of counts of mental disorders in GESIS varied across types of facilities in their level of detail. Among CHD facilities, incidence of mental disorders was aggregated with neurological disorders (coded as “neuro psychiatric disorders”), while CHRR and CHU facilities reported more detailed counts of specific mental disorders, such as depression, personality disorders, and schizophrenia. Across all facilities, disaggregated data was reported by age group but not by sex. Among CSB facilities, incidences of mental disorders were reported by age group under “mental illnesses and psychic disorders” until June 2015, when reporting was reclassified to “mental disorders,” and disaggregated data was reported by both age and sex. A summary of classification by facility type and year is provided in Appendix 2.1, Table S2.1, with English translations of the French disease names provided in Table S2.2. and yearly cumulative reported case totals by disorder reported by facility level and year in Tables S2.3-5.

To address heterogeneity in reporting, our analysis aggregated any monthly incidence of mental or neurological disorders reported by a facility, irrespective of age and sex, into a summary outcome measure. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator[@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation.[@hyndmanAutomaticTimeSeries2008] By considering outliers on a facility-by-facility basis, the data cleaning approach preserved extreme values that may reflect true variances in incidence (for example, incidences reported by a large, regional reference hospital) while addressing potential errors in data entry. (Additional detail on the data cleaning process is available in Appendix 2.1)

@@

ERA 5 is a global atmospheric reanalysis from the European Centre for Medium-Range Weather Forecasts (ECMWF) and is freely available through its Copernicus Climate Change Service (C3S)[@europeanspaceagencyEuropeanSpaceAgency2022]. The spatial resolution of ERA 5 is approximately 30 km, with quality-assured hourly estimates available from 1959 to the present.[@europeanspaceagencyEuropeanSpaceAgency2022] Using Google Earth Engine, a time series of monthly aggregate values for mean air temperature in degrees Kelvin at a height of 2 meters was obtained for each facility geolocation  [@googleearthengineERA5MonthlyAggregates2022] Temperature data was then converted to degrees Celsius.

@@

Duration of cyclone activity was included in all models using our count indicator of exposure. In all models, we included fixed effects for Madagascar’s twenty-two regions to account for geographic differences in climate patterns and case reporting. We further included month and year fixed effects to account for seasonality and time-varying omitted variables.

@@

#### Ethics statement
The Institutional Review Board (IRB) of the Harvard T.H. Chan School of Public Health determined that the study was not human subjects research, and that additional review was not required (protocol number: IRB21-1303, determined on October 19, 2021).

@@

```{r figure_gps, include = TRUE, echo=FALSE, warnings = FALSE,  fig.scap="Map of facilities in Madagascar, by GPS coordinate source and facility type.", out.extra='', fig.cap="\\label{figure_gps}Map of facilities in Madagascar, by GPS coordinate source (left panel) and facility type (right panel).", fig.width = 16, fig.height = 10, fig.align='center', out.height= '90%'}

@@

  # geom_sf(data = map_mdg_3_un,
  #         size = 0.2,
  #         color = col_grey_light,
  #         fill = NA) +
  geom_sf(data = map_mdg_2_un,
          size = 0.2,
          color = col_grey_med,
          fill = NA) +
  geom_sf(data = map_mdg_1_un,
          size = 0.2,
          color = col_black,
          fill = NA) +
  geom_sf(alpha = 0.5, aes(color = source)) +
  theme_void() +
  theme(text = element_text(size = 16, family = "Source Sans Pro"),
        legend.position = "left") +
 scale_color_manual(values = c(col_centroid,
                                col_exact_all,
                                col_exact_rhino,
                                col_exact_val,
                                col_search),
                     name = "GPS type")

@@

model_outputs_preferred_model %>%
  filter(model_type == "NB") %>%
  filter(coefficient %in% c("exposure_temperature:exposure_sm", "exposure_temperature", "exposure_sm", "days_with_cyclone")) %>%
  mutate(
    cell_coef =
      paste0(
        signif(exp(estimate), 3),
        " (",
        signif(exp(lower_99), 3),
        " – ",
        signif(exp(upper_99), 3),
        ")",
        function_pval_stars(p_value)
      ),
    cell_me = ifelse(
      coefficient == "exposure_temperature:exposure_sm",
      "",
      paste0(
        signif(ame, 3),
        " (",
        signif(ame_lower_99, 3),
        " – ",
        signif(ame_upper_99, 3),
        ")",
        function_pval_stars(ame_p_value)
      )
    ),
    coefficient = ifelse(
      coefficient == "exposure_sm",
      "Soil moisture",
      ifelse(
        coefficient == "exposure_temperature",
        "Temperature",
        ifelse(coefficient == "days_with_cyclone",
               "Cyclone (count)",
               "Temperature X Soil moisture"
        )
      )
    ),
    coefficient = factor(
      coefficient ,
      # Reordering group factor levels
      levels = c("Temperature",
                 "Soil moisture",
                 "Temperature X Soil moisture",
                 "Cyclone (count)")
    ),
    model = ifelse(
      model_form == "Temperature only",
      "Model 1 ",
      ifelse(
        model_form == "Soil moisture only",
        "Model 2 ",
        ifelse(model_form == "Temperature + Soil moisture",
               "Model 3 ",
               "Model 4 ")
      )
    ),
    model = factor(model ,      # Reordering group factor levels
                   levels = c("Model 1 ",
                              "Model 2 ",
                              "Model 3 ",
                              "Model 4 "))
  ) %>%
  dplyr::select(model, coefficient, cell_coef, cell_me) %>%
  rename('IRR (99% CI)' = cell_coef,
         'M.E. (99% CI)' = cell_me) %>%
  gt(rowname_col = "coefficient", groupname_col = "model") %>%
  tab_header(title = md("Table 2.2: incidence rate ratios (IRRs) and average marginal effects (AME) of climate exposures on monthly reported cases of mental and neurological disorders, Models 1 through 4.")) %>%

@@

The regression models predicting monthly counts of any mental or neurological disorder (Table 2.2) showed significant associations with temperature across all models. A one-degree Celsius increase in temperature was associated with approximately `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` cases reported `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)` by primary clinics and hospitals in Model 1. Inclusion of both climate predictors resulted in little change to the measure of association for temperature. In an additive model with soil moisture included (Model 3), the effect estimate associated with temperature grew to `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` cases reported `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`. Adopting an interactive model (Model 4) attenuated the association `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`, but remained significant at a 95% level of confidence. Across all models with temperature, the average marginal effect associated with temperature (across a sample with all clinics) ranged between 2 and 3 fewer monthly cases reported and was statistically significant at an alpha of 0.01, with the average marginal effect in Model 4 being `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include ="only")` cases `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include =FALSE)`.

Effect estimates associated with soil moisture were not statistically significant across all four models. In all models, we observed a modest reduction in cases reported per each additional day duration of cyclone activity, with the incidence ratio ratios of Models 3 `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "days_with_cyclone", value = "irr", value_include =TRUE)` and Models 4 `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "days_with_cyclone", value = "irr", value_include = TRUE)` being significant at a level of 0.05. In all models, however, the average marginal effect was close to 0, with the average marginal effect in Model 4 being only `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "days_with_cyclone", value = "ame", value_include ="only")` cases `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "days_with_cyclone", value = "ame", value_include ="FALSE")`.

```{r figure_marginal, include = TRUE, echo=FALSE, warnings = FALSE, fig.width = 13, fig.height = 13, fig.align='center', out.width = '95%', fig.scap="Predicted values of monthly case counts and marginal effect estimates by coefficient and their 95\\% and 99\\% confidence intervals.", out.extra='', fig.cap="\\label{figure_marginal}Predicted values of monthly case counts and marginal effect estimates by coefficient and their 95\\% and 99\\% confidence intervals. The left panels show predicted values of monthly case counts of mental and neurological disorders in Madagascar, given different units of climate exposure. The right panels show marginal effect estimates over different units of climate exposure in the data. The dotted horizontal lines in the right panel depict no marginal effect on case counts. In all panels, the model used to generate predictions and marginal effects corresponds to Model 4. Units of climate exposure are degrees Celsius for temperature (top), saturation percentage for soil moisture (middle), and days with cyclone activity (bottom).", }

panel_l <- outputs_marginal_effects_predictions %>%
  pivot_wider(names_from = level, values_from = c(lower, upper)) %>%
  mutate(factor = ifelse(factor == "exposure_sm", "Soil moisture",
                         ifelse(factor == "exposure_temperature", "Temperature",
                                ifelse(factor == "days_with_cyclone", "Days with cyclone activity",
                                       "ERROR")))) %>%
  mutate(factor = factor(factor ,      # Reordering group factor levels
                         levels = c("Temperature",
                                    "Soil moisture",
                                    "Days with cyclone activity"))) %>%
  ggplot(aes(x = xvals,
             y = yvals,
             group = factor,
             fill = factor,
             color= factor)) +
  # geom_hline(yintercept = 0, color = col_grey_dark, linetype='dotted') +
  geom_ribbon(aes(ymin = lower_0.99,
                  ymax = upper_0.99),
              alpha = 0.15) +
  geom_ribbon(aes(ymin = lower_0.95,
                  ymax = upper_0.95),
              alpha = 0.25) +
  geom_line() +
  geom_point() +
  labs(y = "Predicted monthly case counts, given climate exposure",
       x = "Units of climate exposure") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = col_grey_lightest),
        legend.position = "none",
        strip.background = element_rect(colour="white",
                                        fill="white"),
        strip.text = element_text(size=12),
        text = element_text(family = "Source Sans Pro")) +
  scale_color_manual(values =
                       c(col_temperature,
                         col_soilmoisture,
                         col_cyclone)) +
  scale_fill_manual(values =
                      c(col_temperature,
                        col_soilmoisture,
                        col_cyclone)) +
  facet_wrap(~factor,  scales = "free", nrow = 3)

@@

```{r figure_lag, include = TRUE, echo=FALSE, warnings = FALSE,  fig.scap="Effect estimates by coefficient with lagged exposures and their 99\\% and 95\\% confidence intervals, Model 4.", out.extra='',  fig.cap="\\label{figure_lag}Effect estimates by coefficient with lagged exposures and their 99\\% and 95\\% confidence intervals, Model 4. The x-axis correspond to the length of the lag (e.g., a value of 2 reflects a lag of two months, such that case counts are predicted by climate conditions two months prior), with the value of 0 reflecting a model without any lag in exposure. The dotted horizontal lines depict a null effect. IRR: incidence rate ratio.", fig.width = 8, fig.height = 9, fig.align='center', out.width = '90%'}

@@

```
Predicted values of monthly case counts under Model 4 echoed the effect estimates from Table 2, with fewer cases reported in warmer months (Figure \ref{figure_marginal}). Marginal effect estimates across the distribution of temperatures in our data showed that the greatest level of reduction occurred between 10 and 15 degrees Celsius, with each degree increase associated with 4 to 5 fewer cases of mental or neurological disorders reported, controlling for other covariates. Months with greater levels of soil moisture and days with cyclone activity also showed fewer reported cases; however, the marginal effects associated with soil moisture were not significant across any unit of exposure, while those associated with days with cyclone activity were close to zero for all lengths of activity.

>>>>

     val <- round(exp(output_df$estimate), 3)
     p_val <- round(output_df$p_value, 3)
     low <- format(round(exp(output_df$lower_99), 3), nsmall = 3)
     high <- format(round(exp(output_df$upper_99), 3), nsmall = 3)

@@

     return(paste0(
      "(IRR: ",
       format(val, nsmall = 3),
      ", p ",
      ifelse(p_val < 0.001,
          "< 0.0001, ",
          paste0("= ",
             format(p_val, nsmall = 3),
             ", ")),
      "99\\% CI: ",
      low,
      " – ",
      high
      ,
      ")"))

@@

     return(paste0(
      "(p ",
      ifelse(p_val < 0.001,
          "< 0.0001, ",
          paste0("= ",
             format(p_val, nsmall = 3),
             ", ")),
      "99\\% CI: ",
      low,
      " – ",
      high
      ,
      ")"))

@@

     return(paste0(ifelse(
      val < 1,
      paste0(round(val, 3) * 100, "\\% fewer"),
      paste0(round(val - 1, 3) * 100, "\\%")
     )))

@@

   if (value == "ame"){

    val <- round(output_df$ame, 3)
    p_val <- round(output_df$ame_p_value, 3)
    low <- format(round(output_df$ame_lower_99, 3), nsmall = 3)
    high <- format(round(output_df$ame_upper_99, 3), nsmall = 3)

    if (value_include == TRUE){
     return(paste0(
      "(AME: ",
      format(val, nsmall = 3),
      ", p ",
      ifelse(p_val < 0.001,
          "< 0.001, ",
          paste0("= ",
             format(p_val, nsmall = 3),
             ", ")),
      "99\\% CI: ",
      low,
      " – ",
      high
      ,
      ")"))

@@

     return(paste0(
      "(p ",
      ifelse(p_val < 0.001,
          "< 0.001, ",
          paste0("= ",
             format(p_val, nsmall = 3),
             ", ")),
      "99\\% CI: ",
      low,
      " – ",
      high
      ,
      ")"))

@@

     return(format(round(val, 3), nsmall = 3))

@@

\emph{Daniel Arias, Karestan Koenen, and Christopher Golden}

@@

{\bf Methods:} To better understand climate conditions and population mental health in these settings, we leverage meteorologic and geospatial data to explore the mental health impacts of three CCEs on monthly counts of patient visits for mental and neurological disorders reported by `r num_facilities_pretty` facilities in Madagascar from 2010 to 2020. We used negative binomial regression to examine the impacts of three CCEs: temperature, soil moisture, and duration of cyclones and tropical storms.

{\bf Findings and interpretation:} We found a statistically significant reduction in monthly visits reported associated with temperature—approximately `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = "only")` visits reported per each degree Celsius increase `r function_dynamic_text_results(section = "main", coefficient_option = "exposure_temperature", model_option = 4, value = "irr", value_include = TRUE)`. A multiverse analysis of `r num_models_multiverse` models found this estimate to be robust to alternative specifications; ecoregion analysis indicated that this association was primarily localized to facilities in the central highlands, where cooler temperatures are more prevalent. We further found that higher soil moisture may be associated with higher mental health related visits in the future, with an `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include = "only")` elevation in visits reported `r function_dynamic_text_results(section = "lags", lag_option = 3, coefficient_option = "exposure_sm", value = "irr", value_include =TRUE)` after a three month lag. Ecoregion analysis suggests more visits are reported in the eastern lowlands and the northern dry deciduous forests when soil moisture rises, which is consistent with the geographic concentration of flood risk in Madagascar. Our findings highlight the complex interactions between CCEs and population mental health and suggest potential heterogeneity in how changing climate conditions will impact mental health needs and system responses.

@@

Understanding how both incremental and extreme changes in climate impact population mental health is critical to gauging the full extent to which CCEs impact human health and well-being under a multidisciplinary approach to population health.[@berryCaseSystemsThinking2018] This understanding is particularly important in contexts such as Madagascar, where climate conditions are already demonstrating pronounced pressures on economic, social, and physical well-being.

In investigating the impact of climate exposures on mental health, health system data can serve as a useful tool in settings where epidemiological survey data is sparse. Despite the poor availability of mental health services in many settings, routine health system data can still provide valuable insights into trends and patterns of service use and the need and distribution of care for mental health issues. Although these data may only capture a share of the true burden of disease, reported visits related to mental health complaints can be a valuable starting point for preliminary research into the impact of climate change on population mental health in resource-constrained settings.

@@

As a measure of population mental health, we considered facility-level monthly reported visits related to mental disorders; in terms of exposure, we explored the predictive association of three CCEs with incident mental health visits: mean monthly ambient air temperature, mean monthly soil moisture, and monthly cyclone activity. As exposure to these stressors could involve a delay in onset of mental health symptoms and subsequent care seeking, we further explored whether exposures in prior months were associated with monthly reported visits in subsequent ones. We further considered interactive and additive models to account for complexity in the interaction of climate variables in Madagascar, as well as ecological variation to capture differences in regional climate systems.

@@

To analyze the impact of climate on population mental health in Madagascar, we first obtained health management information system (HMIS) data through a collaboration with the Madagascar Ministère de la Santé Publique. We then obtained gridded temperature and soil moisture data from the European ReAnalysis (ERA5)[@copernicusclimatechangeservicec3sFifthGenerationECMWF2017] and the European Space Agency’s Climate Change Initiative (ESA CCI),[@europeanspaceagencyEuropeanSpaceAgency2022] respectively. Data on tropical storms and cyclones were obtained from open-source compilations of storms in the South-West Indian Ocean[@wikipediaSouthWestIndianOcean2023], which were in turn cross-referenced against storm trajectory data from the National Oceanic and Atmospheric Administration (NOAA)[@internationalbesttrackarchiveforclimatestewardshipibtracsInternationalBestTrack2021] for accuracy.

@@

Exploratory data analysis indicated that the classification and reporting of counts of mental disorders in GESIS varied across types of facilities in their level of detail. Among CHD facilities, incidence of mental disorders was aggregated with neurological disorders (coded as “neuro psychiatric disorders”), while CHRR and CHU facilities reported more detailed counts of specific mental disorders, such as depression, personality disorders, and schizophrenia. Across all facilities, disaggregated data was reported by age group but not by sex. Among CSB facilities, incidences of mental disorders were reported by age group under “mental illnesses and psychic disorders” until June 2015, when reporting was reclassified to “mental disorders,” and disaggregated data was reported by both age and sex. A summary of classification by facility type and year is provided in Appendix 2.1, Table S2.1, with English translations of the French disease names provided in Table S2.2. and yearly cumulative reported visit totals by disorder reported by facility level and year in Tables S2.3-5.

To address heterogeneity in reporting, our analysis aggregated any monthly visits concerning mental or neurological disorders reported by a facility, irrespective of age and sex, into a summary outcome measure. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator[@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation.[@hyndmanAutomaticTimeSeries2008] By considering outliers on a facility-by-facility basis, the data cleaning approach preserved extreme values that may reflect true variances in incidence (for example, incidences reported by a large, regional reference hospital) while addressing potential errors in data entry. (Additional detail on the data cleaning process is available in Appendix 2.1)

@@

ERA5 is a global atmospheric reanalysis from the European Centre for Medium-Range Weather Forecasts (ECMWF) and is freely available through its Copernicus Climate Change Service (C3S)[@europeanspaceagencyEuropeanSpaceAgency2022]. The spatial resolution of ERA5 is approximately 30 km, with quality-assured hourly estimates available from 1959 to the present.[@europeanspaceagencyEuropeanSpaceAgency2022] Using Google Earth Engine, a time series of monthly aggregate values for mean air temperature in degrees Kelvin at a height of 2 meters was obtained for each facility geolocation.[@googleearthengineERA5MonthlyAggregates2022] Temperature data was then converted to degrees Celsius.

@@

Duration of cyclone activity was included in all models using our count indicator of exposure. In all models, we included fixed effects for Madagascar’s twenty-two regions to account for geographic differences in climate patterns and case/visit reporting. We further included month and year fixed effects to account for seasonality and time-varying omitted variables.

@@

#### Ethics statement
The Institutional Review Board (IRB) of the Harvard T.H. Chan School of Public Health determined that this study was not human subjects research and that additional review was not required (protocol number: IRB21-1303, determined on October 19, 2021).

@@

```{r figure_gps, include = TRUE, echo=FALSE, warnings = FALSE, fig.scap="Map of facilities in Madagascar, by coordinate source and facility type.", out.extra='', fig.cap="\\label{figure_gps}Map of facilities in Madagascar, by coordinate source (left panel) and facility type (right panel).", fig.width = 16, fig.height = 10, fig.align='center', out.height= '90%'}

@@

 # geom_sf(data = map_mdg_3_un,
 #     size = 0.2,
 #     color = col_grey_light,
 #     fill = NA) +
 geom_sf(data = map_mdg_2_un,
     size = 0.2,
     color = col_grey_med,
     fill = NA) +
 geom_sf(data = map_mdg_1_un,
     size = 0.2,
     color = col_black,
     fill = NA) +
 geom_sf(alpha = 0.5, aes(color = source)) +
 theme_void() +
 theme(text = element_text(size = 16, family = "Source Sans Pro"),
    legend.position = "left") +
 scale_color_manual(values = c(col_centroid,
                col_exact_all,
                col_exact_rhino,
                col_exact_val,
                col_search),
           name = "Source")

@@

model_outputs_preferred_model %>%
 filter(model_type == "NB") %>%
 filter(coefficient %in% c("exposure_temperature:exposure_sm", "exposure_temperature", "exposure_sm", "days_with_cyclone")) %>%
 mutate(
  cell_coef =
   paste0(
    signif(exp(estimate), 3),
    " (",
    signif(exp(lower_99), 3),
    " – ",
    signif(exp(upper_99), 3),
    ")",
    function_pval_stars(p_value)
   ),
  cell_me = ifelse(
   coefficient == "exposure_temperature:exposure_sm",
   "",
   paste0(
    signif(ame, 3),
    " (",
    signif(ame_lower_99, 3),
    " – ",
    signif(ame_upper_99, 3),
    ")",
    function_pval_stars(ame_p_value)
   )
  ),
  coefficient = ifelse(
   coefficient == "exposure_sm",
   "Soil moisture",
   ifelse(
    coefficient == "exposure_temperature",
    "Temperature",
    ifelse(coefficient == "days_with_cyclone",
        "Cyclone (count)",
        "Temperature X Soil moisture"
    )
   )
  ),
  coefficient = factor(
   coefficient ,
   # Reordering group factor levels
   levels = c("Temperature",
         "Soil moisture",
         "Temperature X Soil moisture",
         "Cyclone (count)")
  ),
  model = ifelse(
   model_form == "Temperature only",
   "Model 1 ",
   ifelse(
    model_form == "Soil moisture only",
    "Model 2 ",
    ifelse(model_form == "Temperature + Soil moisture",
        "Model 3 ",
        "Model 4 ")
   )
  ),
  model = factor(model ,   # Reordering group factor levels
          levels = c("Model 1 ",
               "Model 2 ",
               "Model 3 ",
               "Model 4 "))
 ) %>%
 dplyr::select(model, coefficient, cell_coef, cell_me) %>%
 rename('IRR (99% CI)' = cell_coef,
     'M.E. (99% CI)' = cell_me) %>%
 gt(rowname_col = "coefficient", groupname_col = "model") %>%
 tab_header(title = md("Table 2.2: incidence rate ratios (IRRs) and average marginal effects (AME) of climate exposures on monthly reported visits related to mental and neurological disorders, Models 1 through 4.")) %>%

@@

The regression models predicting monthly counts of any mental or neurological disorder (Table 2.2) showed significant associations with temperature across all models. A one-degree Celsius increase in temperature was associated with approximately `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` visits reported `r function_dynamic_text_results(section = "main", model_option = 1, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)` by primary clinics and hospitals in Model 1. Inclusion of both climate predictors resulted in little change to the measure of association for temperature. In an additive model with soil moisture included (Model 3), the effect estimate associated with temperature grew to `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include ="only")` visits reported `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`. Adopting an interactive model (Model 4) attenuated the association `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "irr", value_include =TRUE)`, but remained significant at a 95% level of confidence. Across all models with temperature, the average marginal effect associated with temperature (across a sample with all clinics) ranged between 2 and 3 fewer monthly mental health visits reported and was statistically significant at an alpha of 0.01, with the average marginal effect in Model 4 being `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include ="only")` visits `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "exposure_temperature", value = "ame", value_include =FALSE)`.

Effect estimates associated with soil moisture were not statistically significant across all four models. In all models, we observed a modest reduction in visits reported per each additional day duration of cyclone activity, with the incidence ratio ratios of Models 3 `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "days_with_cyclone", value = "irr", value_include =TRUE)` and Models 4 `r function_dynamic_text_results(section = "main", model_option = 3, coefficient_option = "days_with_cyclone", value = "irr", value_include = TRUE)` being significant at a level of 0.05. In all models, however, the average marginal effect was close to 0, with the average marginal effect in Model 4 being only `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "days_with_cyclone", value = "ame", value_include ="only")` visits `r function_dynamic_text_results(section = "main", model_option = 4, coefficient_option = "days_with_cyclone", value = "ame", value_include ="FALSE")`.

```{r figure_marginal, include = TRUE, echo=FALSE, warnings = FALSE, fig.width = 13, fig.height = 13, fig.align='center', out.width = '95%', fig.scap="Predicted values of monthly mental health visits and marginal effect estimates by coefficient and their 95\\% and 99\\% confidence intervals.", out.extra='', fig.cap="\\label{figure_marginal} Predicted values of monthly mental health visits and marginal effect estimates by coefficient and their 95\\% and 99\\% confidence intervals. The left panels show predicted values of monthly counts of visits related to mental and neurological disorders in Madagascar, given different units of climate exposure. The right panels show marginal effect estimates over different units of climate exposure in the data. The dotted horizontal lines in the right panel depict no marginal effect on visit counts. In all panels, the model used to generate predictions and marginal effects corresponds to Model 4. Units of climate exposure are degrees Celsius for temperature (top), saturation percentage for soil moisture (middle), and days with cyclone activity (bottom).", }

panel_l <- outputs_marginal_effects_predictions %>%
 pivot_wider(names_from = level, values_from = c(lower, upper)) %>%
 mutate(factor = ifelse(factor == "exposure_sm", "Soil moisture",
             ifelse(factor == "exposure_temperature", "Temperature",
                ifelse(factor == "days_with_cyclone", "Days with cyclone activity",
                    "ERROR")))) %>%
 mutate(factor = factor(factor ,   # Reordering group factor levels
             levels = c("Temperature",
                  "Soil moisture",
                  "Days with cyclone activity"))) %>%
 ggplot(aes(x = xvals,
       y = yvals,
       group = factor,
       fill = factor,
       color= factor)) +
 # geom_hline(yintercept = 0, color = col_grey_dark, linetype='dotted') +
 geom_ribbon(aes(ymin = lower_0.99,
         ymax = upper_0.99),
       alpha = 0.15) +
 geom_ribbon(aes(ymin = lower_0.95,
         ymax = upper_0.95),
       alpha = 0.25) +
 geom_line() +
 geom_point() +
 labs(y = "Predicted monthly mental health visit counts, given climate exposure",
    x = "Units of climate exposure") +
 theme_classic() +
 theme(panel.grid.major.y = element_line(colour = col_grey_lightest),
    legend.position = "none",
    strip.background = element_rect(colour="white",
                    fill="white"),
    strip.text = element_text(size=12),
    text = element_text(family = "Source Sans Pro")) +
 scale_color_manual(values =
            c(col_temperature,
             col_soilmoisture,
             col_cyclone)) +
 scale_fill_manual(values =
           c(col_temperature,
            col_soilmoisture,
            col_cyclone)) +
 facet_wrap(~factor, scales = "free", nrow = 3)

@@

```{r figure_lag, include = TRUE, echo=FALSE, warnings = FALSE, fig.scap="Effect estimates by coefficient with lagged exposures and their 95\\% and 99\\% confidence intervals, Model 4.", out.extra='', fig.cap="\\label{figure_lag}Effect estimates by coefficient with lagged exposures and their 95\\% and 99\\% confidence intervals, Model 4. The x-axis correspond to the length of the lag (e.g., a value of 2 reflects a lag of two months, such that visit counts are predicted by climate conditions two months prior), with the value of 0 reflecting a model without any lag in exposure. The dotted horizontal lines depict a null effect. IRR: incidence rate ratio.", fig.width = 8, fig.height = 9, fig.align='center', out.width = '90%'}

@@

```
Predicted values of monthly mental health related visits counts under Model 4 echoed the effect estimates from Table 2, with fewer visits reported in warmer months (Figure \ref{figure_marginal}). Marginal effect estimates across the distribution of temperatures in our data showed that the greatest level of reduction occurred between 10 and 15 degrees Celsius, with each degree increase associated with 4 to 5 fewer visits related to mental or neurological disorders reported, controlling for other covariates. Months with greater levels of soil moisture and days with cyclone activity also showed fewer reported visits; however, the marginal effects associated with soil moisture were not significant across any unit of exposure, while those associated with days with cyclone activity were close to zero for all lengths of activity.
